---
title: "rrlmgraph Benchmark Report"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rrlmgraph Benchmark Report}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 4,
  out.width  = "100%"
)
library(rrlmgraphbench)
```

## Overview

This vignette presents a reproducible benchmark comparing six retrieval
strategies for providing R coding context to an LLM:

| Strategy | Description |
|---|---|
| `rrlmgraph_tfidf` | rrlmgraph graph traversal with TF-IDF embeddings |
| `rrlmgraph_ollama` | rrlmgraph graph traversal with Ollama embeddings |
| `full_files` | Entire source files dumped verbatim (baseline) |
| `bm25_retrieval` | BM25 keyword retrieval (no graph) |
| `no_context` | No context supplied |
| `random_k` | Five randomly sampled code chunks |

Results are loaded from a pre-generated RDS file to allow the vignette to
knit without an active LLM API key.

---

## Load pre-computed results

```{r load-results, error=TRUE}
results_path <- system.file(
  "results", "benchmark_results.rds",
  package = "rrlmgraphbench"
)

if (file.exists(results_path)) {
  all_results <- readRDS(results_path)
} else {
  message("Pre-computed results not found. Running a dry-run benchmark.")
  all_results <- run_full_benchmark(
    output_path = tempfile(fileext = ".rds"),
    n_trials    = 1L,
    .dry_run    = TRUE
  )
}
str(all_results, max.level = 1)
```

---

## Summary statistics

```{r summary-stats}
stats <- compute_benchmark_statistics(all_results)
knitr::kable(
  stats$summary[, c(
    "strategy", "n", "mean_score", "sd_score",
    "ci_lo_95", "ci_hi_95", "mean_total_tokens", "hallucination_rate"
  )],
  digits = 3,
  caption = "Mean score, 95 % CI, token usage, and hallucination rate per strategy."
)
```

---

## Score distribution

```{r score-plot}
# Base-R dot chart sorted by mean score
summary_df <- stats$summary
summary_df  <- summary_df[order(summary_df$mean_score, decreasing = TRUE), ]

dotchart(
  summary_df$mean_score,
  labels = summary_df$strategy,
  xlab   = "Mean score (0–1)",
  main   = "Strategy performance",
  pch    = 19,
  col    = "steelblue"
)
segments(
  x0   = summary_df$ci_lo_95,
  x1   = summary_df$ci_hi_95,
  y0   = seq_len(nrow(summary_df)),
  lwd  = 2,
  col  = "steelblue"
)
```

---

## Token Efficiency Ratio (TER)

TER measures score-per-token relative to the `full_files` baseline.
A TER > 1 means the strategy achieves a higher score per token consumed.

```{r ter}
ter_df <- data.frame(
  strategy = names(stats$ter),
  TER      = round(stats$ter, 3)
)
ter_df <- ter_df[!is.na(ter_df$TER), ]
ter_df <- ter_df[order(ter_df$TER, decreasing = TRUE), ]
knitr::kable(ter_df, row.names = FALSE,
             caption = "Token Efficiency Ratio relative to full_files baseline.")
```

---

## Hallucination analysis

```{r hallucination}
hall_df <- stats$summary[, c("strategy", "hallucination_rate")]
hall_df$hallucination_rate <- round(hall_df$hallucination_rate, 3)
hall_df <- hall_df[order(hall_df$hallucination_rate), ]
knitr::kable(hall_df, row.names = FALSE,
             caption = "Fraction of responses containing at least one hallucination.")
```

Hallucination type breakdown (where available):

```{r hall-types}
if ("hallucination_details" %in% names(all_results)) {
  details_flat <- unlist(strsplit(
    all_results$hallucination_details[nzchar(all_results$hallucination_details)],
    "; "
  ))
  if (length(details_flat) > 0) {
    type_pattern <- regmatches(
      details_flat,
      regexpr("invented_function|invalid_argument|wrong_namespace", details_flat)
    )
    type_counts <- sort(table(type_pattern), decreasing = TRUE)
    barplot(
      type_counts,
      main  = "Hallucination types",
      ylab  = "Count",
      col   = c("tomato", "goldenrod", "steelblue"),
      names.arg = c("invented\nfunction", "invalid\nargument", "wrong\nnamespace")[
        match(names(type_counts), c("invented_function", "invalid_argument", "wrong_namespace"))
      ]
    )
  } else {
    message("No hallucinations detected in the loaded results.")
  }
}
```

---

## Pairwise statistical tests

Welch *t*-tests with Bonferroni correction and Cohen's *d*.

```{r pairwise}
pw <- stats$pairwise
if (!is.null(pw) && nrow(pw) > 0) {
  pw$sig <- ifelse(pw$p_bonferroni < 0.001, "***",
             ifelse(pw$p_bonferroni < 0.01,  "**",
             ifelse(pw$p_bonferroni < 0.05,  "*", "")))
  knitr::kable(
    pw[, c("strategy_1", "strategy_2", "statistic",
           "p_value_raw", "p_bonferroni", "cohens_d", "sig")],
    digits  = 4,
    caption = "Pairwise Welch t-tests. * p<0.05; ** p<0.01; *** p<0.001 (Bonferroni)."
  )
}
```

---

## Per-project breakdown

```{r project-breakdown}
if ("task_id" %in% names(all_results)) {
  # Infer project from task_id (e.g. task_001_fm_mini_ds → mini_ds)
  all_results$project <- sub(".*_(mini_ds|shiny|rpkg).*", "\\1",
                              all_results$task_id)
  proj_summary <- aggregate(
    score ~ strategy + project,
    data = all_results,
    FUN  = mean
  )
  proj_wide <- reshape(
    proj_summary,
    idvar    = "strategy",
    timevar  = "project",
    direction = "wide"
  )
  names(proj_wide) <- gsub("score\\.", "", names(proj_wide))
  knitr::kable(
    proj_wide,
    digits  = 3,
    caption = "Mean score disaggregated by fixture project."
  )
}
```

---

## Recursive-improvement trajectory

If multi-trial results are available, this plot shows whether score
improves across successive trials (a proxy for benefit from context).

```{r improvement-plot}
if ("trial" %in% names(all_results)) {
  trial_means <- aggregate(score ~ strategy + trial,
                           data = all_results, FUN = mean)
  strategies  <- unique(trial_means$strategy)
  cols        <- rainbow(length(strategies))
  plot(
    range(trial_means$trial), c(0, 1),
    type = "n",
    xlab = "Trial", ylab = "Mean score",
    main = "Score by trial"
  )
  for (i in seq_along(strategies)) {
    sub <- trial_means[trial_means$strategy == strategies[i], ]
    lines(sub$trial, sub$score, col = cols[i], lwd = 2, type = "b", pch = 19)
  }
  legend("bottomright", legend = strategies, col = cols, lwd = 2, cex = 0.8)
}
```

---

## Session info

```{r session-info}
sessionInfo()
```
